### 2021-11-14 dm.tex

% Standalone document
\documentclass[notes.tex]{subfiles}
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Supersymmetric dark matter}
\label{chap:dm}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Evidence for dark matter (DM)}
The history of dark matter goes back quite a long way. Today we have evidence for the existence of dark matter through several effects where we observe its gravitational influence on ordinary matter.  We list the evidence below:
\begin{enumerate}[1)]
\item Kinematics  (Zwicky 1933~\cite{Zwicky:1933gu}): The motion of galaxies (velocity dispersion) cannot be explained by the visible matter. This was also observed on the scales of galaxies in their rotation curves (Rubin 1970~\cite{Rubin:1970zz}).
\item Gravitational lensing (Tyson 1996~\cite{Fischer:1996eb}). First observed in galactic clusters. Clusters show evidence of lensing not explained by luminous matter.  Dark matter dynamics (non-interacting) are demonstrated by the Bullet cluster (Clowe 2006~\cite{Clowe:2006eq}). 
\item Large scale structures (clusters, superclusters, filaments and voids): The 2dFGRS (2-degree Field Galaxy Redshift survey~Colles 2001~\cite{Hawkins:2002sg}) and SDSS (Sloan Digital Sky Survey~Tegmark 2004~\cite{Abazajian:2004it}) give a relative matter density of $\Omega_m \equiv \frac{\rho_m}{\rho_c} = 0.29$ where $\rho_c = 1.05\cdot 10^{-5}h^2$\,GeV/cm$^3$ is the critical energy density for a flat universe.\footnote{$h$ is defined through the Hubble constant $H_0$ as $H_0=100 h$\,km/Mpc/s.} They also imply that the majority of DM must be {\bf cold} (non-relativistic), because warm DM would suppress clustering.
\item Big-Bang Nucleosynthesis (BBN): The formation of light elements in the period $t = 1-1000$\,s after the Big Bang. Measurements of Early Universe abundance of light elements, mainly D and He, points to a baryonic matter density of $\Omega_b \approx 0.04$. This gives $\Omega_{\text{leftover}} \approx 0.25$.
\item Supernovae (Riess 1998~\cite{Riess:1998cb} and Perlmutter 1999~\cite{Perlmutter:1999jt}): Measurements of type Ia supernovae (SNe Ia) were used as standard candles to show an accelerated expansion of the Universe. This fixes $\Omega_\Lambda-\Omega_m \simeq k$ where $\Omega_\Lambda$ is the energy density of dark energy/cosmological constant.
\item Cosmic Microwave Background (CMB) (Penzias \& Wilson 1965~\cite{Penzias:1965wn}): The temperature variation of the CMB over the sky of the order of $0.0002$\,K is sensitive to all cosmological parameters, and gives $\Omega_\Lambda +\Omega_m \simeq k$, where $k$ is some constant.
\end{enumerate}

The evidence above can be used to constrain a minimal model of the Universe that can explain all the current measurements, the $\Lambda CDM$ concordance model of cosmology, which has just a handful of ingredients such as baryonic and dark matter, radiation (photons) and dark energy.  In Fig.~\ref{fig:LambdaCDMfit} we show the effects of the SNe, CMB and large scale structure data (BAO) on this model.

\begin{figure}[h!] 
\includegraphics[width=0.5\textwidth]{figures/DMfit.eps} 
\caption{Limits from different experiments on the dark energy density, $\Omega_\Lambda$, compared to the total mass density in the universe, $\Omega_m$. \label{fig:LambdaCDMfit}}
\end{figure}

A maximum likelihood fit to a selected subset of the measurements gives the parameters for the model shown in Table~\ref{tab:LambdaCDM}.

\begin{table}[h!]
\begin{tabular}{l |c| c| c| c| c} 
\hline
Parameter & $\Omega_\Lambda$ & $\Omega_{m} h^2$ & $\Omega_b h^2$ & $H_0$\,[km/Mpc/s] & $t_0$\,[Gy]\\\hline
Value & $0.685^{+0.018}_{-0.016}$& $0.1426\pm 0.0025$ & $0.02205 \pm 0.00028$ & $67.3\pm 1.2$ & $13.817 \pm 0.048$ \\
\end{tabular}
\caption{Measured values for cosmological parameters~\cite{Ade:2013zuv}.}\label{tab:LambdaCDM}
\end{table}

\section{WIMP magic}
The very existence of a stable Weakly\footnote{Weak as in electro-weak, meaning on the same scale as the weak force.} Interacting Massive Particle (WIMP) $\chi$ automatically gives an additional component to the total energy density of the Universe. WIMPs are found in a number of theories, for example the lightest neutralino of the MSSM, the lightest Kaluza-Klein particle of a theory with extra dimensions or an inert Higgs boson.

This is due to the in equilibrium {\bf thermal production} of the WIMP through the process $SM\times SM \to \chi \chi$, and the reverse annihilation process $\chi \chi \to SM\times SM$, in the early hot Universe ($T\gg m_\chi$). As the temperature decreases to $T< m_\chi$ and there is not enough energy in an average collision for the production of $\chi$ to occur, only the reverse process can take place, and the comoving density\footnote{Taking the expansion of the universe into account by looking at the number of particles in a volume expanding at the same rate.} falls with the temperature of the Universe.

The WIMPs then experience what is called a {\bf chemical decoupling}, or loss of chemical equilibrium, due to the expansion of the Universe. This is when the WIMPs become so dillute, because of the expansion, that they in effect no longer interact inelastically, and this roughly happens when the expansion rate becomes larger than the rate of anihilation. The WIMPs then get a constant (comoving) density, we say that they experience a {\bf freeze-out} at this temperature $T_c$. With weak-scale masses and couplings the freeze-out happens at $T_c \approx 0.05 m_{\chi}$, however, this is before or at the same time as {\bf kinetic decoupling} where the WIMPs effectively lose elastic interactions, meaning that $\chi$ freezes-out with non-relativistic velocities and become so-called cold dark matter.

The exact time (temperature) of freeze-out is controlled by the annihilation cross section of $\chi$, larger cross sections keep chemical equilibrium for longer, in turn resulting in lower dark matter relic abundance. This abundance, in number density, can be found from the Boltzmann equation
\begin{equation}
\frac{dn_\chi}{dt} = -3H n_\chi - \langle \sigma v \rangle (n_{\chi}^2 - n_{\chi}^{eq}{}^2),
\end{equation}
where $n_{\chi}^{eq}$ and $n_{\chi}$ are the chemical equilibrium and actual comoving number densities, $H$ is Hubble's constant for the expansion rate, and $\langle \sigma v\rangle$ the velocity averaged annihilation cross section for $\chi\chi \to SM \times SM$. In practice one must also often take into account co-annihilation with other particles with mass within $10\% - 20\%$ of the $\chi$, and numerical codes such as {\tt DarkSUSY}~\cite{Gondolo:2004sc} or {\tt MicrOMEGAs}~\cite{Belanger:2001fz,Belanger:2004yn} are used.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=\textwidth]{figures/freezeout.eps} 
\caption{Ilustration of the freeze-out of the comoving number density of a WIMP, where the black line represents a model without chemical decoupling, and the dotted lines represent different freeze out temperatures for different velocity averaged annihilation cross sections. \label{freeze}}
\end{center}
\end{figure}

For weak scale particles a rough approximation to the resulting dark matter density is 
\[\Omega_\chi h^2 = 0.1 \times \frac{3\times 10^{-26}\,{\rm cm}^3 {\rm s}^{-1}}{\langle \sigma v \rangle},\]
and since the annihilation cross section can be shown to be
\begin{equation}
\langle \sigma v \rangle \approx \frac{\alpha_{weak}^2}{m_{weak}^2} \approx 10^{-25}\,{\rm cm}^3 {\rm s}^{-1},
\end{equation}
the predicted DM density is
\[\Omega_\chi h^2 \approx 0.1 \times \left(\frac{g_{weak}}{g_\chi}\right)^4\left(\frac{m_\chi}{m_{weak}}\right)^2.\]
When compared to the value in Table~\ref{tab:LambdaCDM} this is called the {\bf WIMP-miracle}.

For a more detailed discussion of the WIMP miracle, see the standard cosmology book by Kolb and Turner~\cite{Kolb:1990vq}.

%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dark matter candidates in supersymmetry}
%%%%%%%%%%%%%%%%%%%%%%%%%%
With R-parity conservation in place we have seen that  any neutral LSP can be DM. Without R-parity only super-weakly coupling particles like gravitinos and axinos are candidates. Below we briefly discuss the various possibilities.

\subsection{Neutralino}
As soon as you have a stable neutralino LSP, you usually get into trouble trying to explain why there is so little dark matter. The neutralino in the standard mSUGRA bino like $\tilde{\chi}^0_1$ scenario gives a problematically high $\Omega_\chi h^2$ due to current lower bounds on the $\tilde{\chi}^0_1$ mass and the measured higgs mass. This is called the {\bf bulk region} scenario which can be seen in Fig.~\ref{NeuDM} in the lower left corner.  Alternatives to the bulk region scenario use co-annihilation or resonant annihilation to increase $\langle \sigma v \rangle$ and thus decrease the dark matter density.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=\textwidth]{figures/NeutralinoDM.eps} 
\caption{Generic illustration of the allowed neutralino DM regions (puke green) in the $(m_0,m_{1/2})$-plane for mSUGRA. Except for the low $m_0$ and $m_{1/2}$ regions the area outside of the allowed region gices too much drk matter. The dashed line shows the Higgs mass limit which pushes towards larger values of $m_{1/2}$, while the dotted line represents the limit from the anomalous magnetic moment of the muon.\label{NeuDM}}
\end{center}
\end{figure}

The {\bf stau-coannihilation region}, where $\tilde{\tau}_1 \tilde{\chi}^0_1 \to SM\times SM$ is an efficient process, exists for small $m_0$ with $m_{\tilde{\tau}_1} - m_{\tilde{\chi}^0_1}\leq 10$\,GeV, which makes this scenario difficult to discover at collider experiments due to the production of soft (low-energy) taus. This is  shown as the lower strip in Fig.~\ref{NeuDM} which follows the lower theoretical bound (brown) where the stau becomes the LSP.

The {\bf stop-coannihilation region}, where $\tilde{t}_1 \tilde{\chi}^0_1 \to SM\times SM$, exists for large values of $|A_0|$, small $m_0$ and  $m_{1/2}$, and typically has $m_{\tilde{t}_1} - m_{\tilde{\chi}^0_1}\leq 25$\,GeV. Again, this is difficult to discover because of the soft decay products of the stop.

The {\bf higgs funnel region} for $2 m_{\tilde{\chi}^0_1} = m_{A,H}$ and large $\tan \beta$, where the neutralino has resonant annihilation through a heavy Higgs boson, is shown in Figure \ref{NeuDM} as the diagonal structure roughly in the middle of the plot, rising as a funnel upwards.

The {\bf focus point region} for large $m_0$ and low $\mu$ gives an higgsino-wino LSP with a more efficient higgsino annihilation channel for the LSP and thus a lower dark matter density within experimental bounds. This leads to so-called split-SUSY, as the sfermion masses need to be pressed up quite a bit. The focus point region can be seen in Fig.~\ref{NeuDM} following the upper theoretical bound where EWSB breaks down.

%%%
\subsection{Sneutrinos}
%%%
The left handed sneutrino $\tilde{\nu}_L$ is happily excluded as a DM particle due to the large cross section for  $\tilde{\nu}_L q \to \tilde{\nu}_L q$ via $Z$-exchange.\footnote{For the neutralinos this problem only exists for a higgsino $\tilde{\chi}^0_1$ LSP, as the wino and bino  do not couple to the $Z$} The large cross means that it should already have been seen by direct detection experiments. It is also problematic to get $m_{\tilde{\nu}_L}<m_{\tilde{l}_L}$ due to hyperfine-splitting. However, $\tilde{\nu}_R$ couples very weakly and is still a viable candidate.

%%%
\subsection{Gravitino}
%%%
The gravitino is {\it not} a WIMP  as  it is never in chemical equilibrium. It can be created from NLSP decays giving, in RPC scenarios,
\[\Omega_{\tilde{G}} = \frac{m_{\tilde{G}}}{m_{NLSP}}\Omega_{NLSP},\]
however, these scenarios are problematic, because the NLSP is long-lived  and creates potential trouble in BBN by injecting energy that changes the production of light elements. Alternatively, it can be created in non-thermal production as shown in Fig.~\ref{gravterm} at reheating after inflation. The reverse process $\tilde{g}\tilde{G} \to gg$ is not efficient as the density of gravitinos and gluinos is never high enough given the small the cross section. This type of dark matter creation process is often called {\bf freeze-in}. For the gravitino this gives a new magic formula:
\begin{equation}
\Omega_{\tilde{G}}h^2 \approx 0.5 \cdot \left(\frac{T_R}{10^{10}\,{\rm GeV}}\right)\left(\frac{100\,{\rm GeV}}{m_{\tilde{G}}}\right)\left(\frac{m_{\tilde{g}}}{1\,{\rm TeV}}\right)^2,
\end{equation}
where $T_R$ is the reheating temperature. This is valid also for RPV scenarios. There the gravitino coupling $\propto \frac{1}{M_P}$ makes the gravitinos very long-lived, but not absolutely stable. One can also imagine an axino scenario, that would work just like the gravitino.

\begin{figure}[h!]
\begin{center}
\includegraphics{figures/gravterm.eps} 
\caption{One possible diagram for the non-thermal production of gravitinos.\label{gravterm}}
\end{center}
\end{figure}

\subsection{Others}
One could even imagine color charged supersymmetric particles as DM, in particular the gluino, which, if stable, after hadronization form so-called {\bf R-hadrons}.  These have very strict limits from direct searches, but these limits are somewhat obfuscated by complications in R-hadron scattering.

%%%%%%%%%%%%%
\section{Direct detection}
%%%%%%%%%%%%%
In addition to the direct production of dark matter at colliders and the corresponding searches for missing energy, there are two other main ways to search for dark matter, direct and indirect detection. Here we briefly discuss direct detection.

Direct detection seeks to make weak DM interactions with SM matter visible by very low background searches in large volumes, using galactic halo DM interacting with ordinary matter. This is very dependent on the $\chi$ scattering cross section on nucleons (quarks), which can be calculated in a given model, but also the DM halo density distribution and velocity distribution, which have large uncertainties. This can be expressed in the differential scattering rate with respect to the recoil energy $E_r$ of a scattered nucleon with mass $M$
\begin{equation}
\frac{dN}{dE_r} = \frac{\sigma \rho_{DM}}{2\mu^2 m_\chi}|F(q)|^2\int_{v_{min}}^{v_{esc}} \frac{f(\vec v)}{v}d^3v,
\end{equation}
where $\sigma$ is the DM scattering cross section off the nucleus in question, $\rho_{DM}$ is the DM halo density at Earth, $\mu$ is the dark matter and nucleus reduced mass, $F(q)$ is a nuclear form factor dependent on the scattering momentum transfer $q=\sqrt{2ME_r}$, $f(\vec v)$ is the velocity distribution in the halo, $v_{min}=\sqrt{ME_r/2\mu^2}$ is the minimal velocity that gives a recoil energy $E_r$ and $v_{esc}$ is the escape velocity from the halo.

There two main tactics followed in order to try to directly detect DM:
\begin{itemize}
\item Suppress (almost) all backgrounds, which is used in experiments such as XENON and CDMS.
\item Look for an annular modulation, due to the Earth's movement in the galactic rest frame, in a small dark matter signal on top of a constant background, used in DAMA and CoGeNT.
\end{itemize}

Figure~\ref{DD} shows results from the most important direct detection experiments. Observe that while DAMA and CoGeNT both have signals for detection, these are already excluded by XENON and not compatible with each other.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=\textwidth]{figures/DMexpplot.eps} 
\caption{Plot of different exclusion and detection results for direct detection of DM in the WIMP mass versus WIMP--nucleon cross section plane. The grey area shows the expected mass and cross section in MSSM models, where we assume gauge unification at the GUT-scale. \label{DD}}
\end{center}
\end{figure}

%%%%%%%%%%%%%
\section{Indirect detection}
%%%%%%%%%%%%%
In indirect detection we look for annihilation or decay products from DM in multiple final (messenger) states in cosmic rays. Search channels must be stable SM particles, so that they can reach the Earth (or satelites in orbit). The messengers should also have as low backgrounds from ordinary astrophysical processes as possible, this makes searches with electrons and protons difficult. The remaining candidates are photons, neutrinos, positrons, antiprotons and antideuterons.

\begin{itemize}
\item {\bf Photons:} These can either come from direct production processes such as $\chi\chi \to \gamma \gamma, Z\gamma$, which is easier to detect because the spectrum is a sharp line spectrum at exactly the mass of the DM, or $\gamma$ from brehmsstrahlung or pion decays, which is a broad spectrum and hard to detect, but is expected to make up the majority of photons from dark matter. Photons from dark matter have the advantage that they point to the source so we can focus on areas with large $\rho_{DM}$, and thus reducing potential backgrounds relative tot he signal. We can also look for photons that are extragalactic in origin (but then we have to account for red-shifting the spectrum).

Dark matter annihilating in our own galaxy into photons should result in a flux at Earth given by
\begin{equation}
\frac{d\Phi}{dE d\Omega} = \frac{1}{8\pi m_\chi^2}\frac{dN_\gamma}{dE}\langle \sigma v \rangle \int_{\text{l.o.s.}} \rho^2_{DM}(l) dl,
\end{equation}
where $N_\gamma (E)$ gives the number of photons with energy $E$ in a single annihilation event. We see that the flux depends on the square of the DM density since annihilation requires two DM particles to be present. For decaying DM the corresponding expression is proportional to $\rho_{DM}$.

There have been som indications of an excess of photons above expected backgrounds from the galactic centre (a.k.a. the Hooperon~\cite{Hooper:2010mq}), however, no unambigeous DM signal has been confirmed. Current limits from the Fermi-LAT experiment seems to rule out most possible models a DM explanation for this excess, are sets a cross section limit close, and for some masses beyond, to the canonical limit
\[\langle\sigma v\rangle = 3\times 10^{-26}\,{\rm cm}^{3}{\rm s}^{-1},\]
see Fig.~\ref{fig:FERMI}.

\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/FERMI_dSph_bb.eps} 
\caption[Fermi-LAT results]{Results from Fermi-LAT indirect gamma-ray searches in the $\chi\chi\to b\bar b$ channel. Grey line shows limit from Milky Way halo search, black line from Milky Way dwarf spheroidal galaxy search with six years of data~\cite{Ackermann:2015zua}. \label{fig:FERMI}}
\end{center}
\end{figure}

\item {\bf Neutrinos:} These also point to the source and can be extragalactic in origin just like the photons. The same flux calculation can be used, starting from the neutrino spectrum from dark matter annihilation. The astrophysical background is smaller, however, the neutrino signal is difficult to detect. The current leading experiment is IceCube at the South Pole. One interesting possibility is that DM matter scatters on ordinary matter sufficiently strongly that DM accumulates at the centre of the Sun (or possibly the Earth). When these DM particles annihilate the only decay products that can escape the Sun's interior are neutrinos.

\item {\bf Positrons:} Charged particles propagate in a complicated way through the galactic magnetic field, and they are therefore impossible to track back to the source. Sources outside of our own Galaxy cannot contribute significantly to the flux at Earth. This source has large astrophysical backgrounds, so experiments search for small excesses, mostly at high energies. Some potential excess has been seen by Fermi-LAT and PAMELA.

\item {\bf Antiprotons:} As the positrons these propagate in a complicated manner, but the backgrounds are under better control. PAMELA has set strict limits.

\item {\bf Antideuterons:} These have very low backgrounds because, however, the physics of the formation is quite complicated and hard to calculate. AMS-02 will provide new data soon.
\end{itemize}

%%%%%%%%%%%%%
\section{Excercises}
%%%%%%%%%%%%%

\begin{exercise}
Show that $\chi\chi \to Z \to f \bar{f}$ gives
\begin{equation}
\sigma v \approx \frac{g^4 E^2_\chi}{128 \pi m_Z^2},
\end{equation}
which in the low-velocity limit can be shown to be
\[\langle\sigma v\rangle_0 \approx 10^{-25}\,{\rm cm}^3{\rm s}^{-1}.\]
\end{exercise}

\end{document}